---
title: "robbins_metro_bikeshare_draftfinal"
author: "Thomas Robbins"
date: "2023-11-29"
output: html_document
---

This Rmd file was created at the beginning of my group's final project, where I wanted to set my group up for success by collecting data, tidying the data, and creating preliminary graphics and trend analysis. My work in creating this code practiced key skills in R and critical data science principles while laying the groundwork for our group's project, and therefore I believe it exemplifies my learning from this course:

Please not that a portion of data cleaning shown below was completed with the help of Eric Fingar on my team.

Loading packages:
```{r}
library(tidyverse)
library(janitor)
library(here)
library(rwunderground)
library(openmeteo)
library(ggplot2)
library(geosphere)
#library(read)
```

Loading bike data:
```{r}
thisfile=here("data_raw", "202305-capitalbikeshare-tripdata.zip")

df1=read_csv(thisfile) %>% clean_names()
```
Loading metro station data:
```{r}
thisfile=here("data_raw","Metro_Station_Entrances_in_DC.csv")

dfM=read_csv(thisfile) |> clean_names()
```

Combining start and end data in bikeshare:
```{r}
# Combine start and end data
df2s <- df1 %>%
  select(rideable_type, member_casual, contains("start"), ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at, station_name = start_station_name, station_id = start_station_id, lat = start_lat, lng = start_lng)

df2e <- df1 %>%
  select(ride_id, rideable_type, member_casual, contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at, station_name = end_station_name, station_id = end_station_id, lat = end_lat, lng = end_lng)

df2 <- bind_rows(df2s, df2e) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)

```


Creating a cumulative ridership vs location data frame:
```{r}
# Create a dataframe with total ridership, lat, lng for unique station IDs
df_total_ridership <- df2 %>%
  group_by(station_id, lat, lng) %>%
  summarise(total_ridership = max(riders)) %>%
  ungroup()

```
Locating metro stations:
```{r}

# Create a new dataframe with only lat and long columns
df_metro_coordinates <- data.frame(lat = dfM$y, lng = dfM$x)

```

Cleaning data and merging bike and metro stations based on closest proximity:
```{r}

# Create df_cleaned with total_ridership, latbike, and lngbike
df_cleaned <- df_total_ridership %>%
  group_by(station_id) %>%
  summarise(
    total_ridership = sum(total_ridership),
    latbike = first(lat),  # Assuming lat and lng are the original columns
    lngbike = first(lng)   # Adjust if your columns are named differently
  )

# Add metro station coordinates with swapped column names
df_cleaned <- merge(
  df_cleaned,
  df_metro_coordinates %>% select(lat, lng) %>% rename(latmetro = lat, lngmetro = lng),
  by = character(),
  all.x = TRUE
)

```

Continued...
```{r}
# Assuming df_cleaned has the columns "latbike", "lngbike", "latmetro", "lngmetro"
df_cleaned_with_distance <- df_cleaned %>%
  rowwise() %>%
  mutate(
    dist_to_metro = distHaversine(c(lngbike, latbike), c(lngmetro, latmetro)) / 1609.344
  )

```


Plotting ridership vs distance to metro:
```{r}

ggplot(df_cleaned_with_distance, aes(x = dist_to_metro, y = total_ridership))+
    geom_point(alpha = 0.05, color = "#074650", size = 0.75) +
    geom_smooth(method = "loess", se = FALSE,
                color = "limegreen", linewidth = 1.5) +  
  labs(title = "Relationship between Ridership and Distance to Metro",
       x = "Distance to Metro (miles)",
       y = "Total Ridership" ) +
  xlim(0,25) +
  ylim(0,150000)
  theme_minimal()

```

