---
title: 'STAT383 Final Project: Bikeshare'
output:
  html_document:
    df_print: paged
---

Toward the end of the semester, I used my DS241 skillset to do further analysis of bikeshare, this time for my final STAT383 project doing regression analysis to understand optimal bikeshare conditions in Seoul, South Korea.

The following code was all written by myself, and I feel it exemplifies my capacity to put together organized code, analyze and clean data, use various key functions in R, and produce effective visualizations.

Referencing key packages:
```{r}

library(tidyverse)
library(janitor)
library(here)
library(rwunderground)
library(openmeteo)
library(ggplot2)
library(geosphere)

```

Loading raw Seoul bikeshare date (hourly):
```{r}

thisfile=here("data_raw", "SeoulBikeData.csv")

bikes_raw = read_csv(thisfile) %>% clean_names()

```

Removing outlier days (holidays and nonfunctioning), removing variables not relevant to regression, and combining precipitation variables by equivalent rain height (NSSL 1mm rain = 13mm snow) (eql_rain_mm):
```{r}

bikes_filter = bikes_raw |> 
  filter(holiday == "No Holiday", functioning_day == "Yes") |> 
  select(-date, -dew_point_temperature, -solar_radiation_mj_m2, -seasons, -holiday, -functioning_day)

bikes_total = bikes_filter |> 
  mutate(eql_rain_mm = rainfall_mm + snowfall_cm/10/13) |> 
  select(-rainfall_mm, -snowfall_cm)

glimpse(bikes_total)

```

Building linear regression model for ridership as a function of the following variables- hour, temperature, humidity, windspeed, visibility, and precipitation:
```{r}

regression_total = lm(data = bikes_total, rented_bike_count ~ hour + temperature_c + humidity_percent + wind_speed_m_s + visibility_10m + eql_rain_mm)

summary(regression_total) 

# This builds a regression model to predict total hourly ridership as with all variables as inputs, as opposed to drawing intependent trends for each variable.

```

Independent regression model for ridership vs hour of day--

Key output values: 
 - Coefficient/slope: 40.1596
 - Intercept: 277.1531
 - Std Error of Slope: 0.9353
 - t Value: 42.94
 - Adj R Squared: 0.1862
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 581 on 8055 deg of freedom
 
```{r}

reg_hour = lm(data = bikes_total, rented_bike_count ~ hour)
summary(reg_hour) 

hour_plot = ggplot(data = bikes_total,
                   aes(x = hour, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 277.1531,
                        slope = 40.1596, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Hour of Day",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Hour of Day") +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
hour_plot

```

Independent regression model for ridership vs temperature--

Key output values: 
 - Coefficient/slope: 29.4352
 - Intercept: 358.7187
 - Std Error of Slope: 0.4927
 - t Value: 59.74
 - Adj R Squared: 0.307
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 536.1 on 8055 deg of freedom
 
```{r}

reg_temp = lm(data = bikes_total, rented_bike_count ~ temperature_c)
summary(reg_temp) 

temp_plot = ggplot(data = bikes_total,
                   aes(x = temperature_c, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 358.7187,
                        slope = 29.435, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Hourly Temperature (Deg C)",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Temperature") +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
temp_plot

```

Independent regression model for ridership vs humidity--

Key output values: 
 - Coefficient/slope: -6.7230
 - Intercept: 1131.6885
 - Std Error of Slope: 0.3435
 - t Value: -19.57
 - Adj R Squared: 0.04527
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 629.3 on 8055 deg of freedom
 
```{r}

reg_hum = lm(data = bikes_total, rented_bike_count ~ humidity_percent)
summary(reg_hum) 

hum_plot = ggplot(data = bikes_total,
                   aes(x = humidity_percent, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 1131.6885,
                        slope = -6.7230, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Hourly Percent Humidity",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Humidity") +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
hum_plot

```

Independent regression model for ridership vs windspeed-- 

Key output values: 
 - Coefficient/slope: 85.106
 - Intercept: 593.025
 - Std Error of Slope: 6.926
 - t Value: 12.29
 - Adj R Squared: 0.01828
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 638.1 on 8055 deg of freedom
 
```{r}

reg_wind = lm(data = bikes_total, rented_bike_count ~ wind_speed_m_s)
summary(reg_wind) 

wind_plot = ggplot(data = bikes_total,
                   aes(x = wind_speed_m_s, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 593.025,
                        slope = 85.106, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Hourly Wind Speed (m/s)",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Wind Speed") +
  xlim(0,5) +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
wind_plot

```

Independent regression model for ridership vs visibility-- 

Key output values: 
 - Coefficient/slope: 2.352e-02
 - Intercept: 4.028e+02
 - Std Error of Slope: 1.153e-03
 - t Value: 20.41
 - Adj R Squared: 0.04904
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 628 on 8055 deg of freedom
 
```{r}
bikes_total_2 = bikes_total |> 
  mutate(visibility_m = visibility_10m * 10) |> 
  select(-visibility_10m)

reg_vis = lm(data = bikes_total_2, rented_bike_count ~ visibility_m)
summary(reg_vis) 

vis_plot = ggplot(data = bikes_total_2,
                   aes(x = visibility_m, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 4.028e+02,
                        slope = 2.352e-02, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Visibility (meters)",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Visibility") +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
vis_plot

```

Independent regression model for ridership vs precipitation--

Key output values: 
 - Coefficient/slope: -73.849
 - Intercept: 750.591
 - Std Error of Slope: 6.213
 - t Value: -11.89
 - Adj R Squared: 0.01711
 - p Value (F statistic): < 2.2e-16
 - Res Std Err: 638.5 on 8055 deg of freedom
 
```{r}

reg_prec = lm(data = bikes_total, rented_bike_count ~ eql_rain_mm)
summary(reg_prec) 

prec_plot = ggplot(data = bikes_total,
                   aes(x = eql_rain_mm, y = rented_bike_count)) +
            geom_point(alpha = 0.1, color = "darkgrey") +
            geom_smooth(aes(color = "Smooth Average Correlation Curve"), linewidth = 1.5) +
            geom_abline(aes(intercept = 750.591,
                        slope = -73.849, color = "Linear Regression Model Output"),
                        linetype = "dashed",linewidth=1.5) +
            theme_minimal() +
            labs(x = "Precipitation (mm equivalent rainfall)",
                 y = "Hourly Total Count of Rented Bikes",
                 title = "Ridership vs Precipitation") +
  xlim(0,5) +
  ylim(0,3000) +
  scale_color_manual(values = c("Smooth Average Correlation Curve" = "green",
                                "Linear Regression Model Output" = "red" )) +
  guides(color = guide_legend(title = "Linear Annotations")) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))
                   
prec_plot

```



